---
title: "Trends in Graduate Enrollment, Shifting Fields of Study, and Faculty Salary"
author: "Sam Csik"
date: "11/28/2018"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r load packages & data}

library(tidyverse)
library(reshape2) 
library(effsize)
library(scales)
library(extrafont)
library(car)
library(stargazer)
library(beyonce)

enrollment <- read_csv("grad_enrollment_copy.csv")
postdoc_salary <- read_csv("median_salary_copy.csv")
fields <- read_csv("phds_by_field_copy.csv")
faculty_salary <- read_csv("faculty_salary.csv")
phd_fields_field <- read_csv("phds_field_field.csv")

```

```{r male vs. female grad enrollment (1967-2015)}

# SOURCE: U.S. Department of Education, National Center for Education Statistics, Higher Education General Information Survey (HEGIS), “Fall Enrollment in Colleges and Universities” surveys, 1967 through 1985; Integrated Postsecondary Education Data System (IPEDS),“Fall Enrollment Survey” (IPEDS-EF:86-99); IPEDS Spring 2001 through Spring 2016, Fall Enrollment component. Accessed at: https:// nces.ed.gov/ programs/digest/ d16/ tables/ dt16_303.80.asp.

##############################
# subset total males and females
##############################

total_males <- enrollment %>% 
  select(year, total_males) %>% 
  mutate(total = total_males) %>% # add column called "total" that is equal to the "total_males" column so that we can plot males and females on same plot with same y-axis 
  mutate(sex = rep("male", length(total_males))) %>% 
  select(year, total, sex)

total_females <- enrollment %>% 
  select(year, total_females) %>% 
  mutate(total = total_females) %>%  # add column called "total" that is equal to the "total_females" column so that we can plot males and females on same plot with same y-axis
  mutate(sex = rep("female", length(total_females))) %>% 
  select(year, total, sex)

total <- rbind(total_males, total_females)

##############################
# plot trends through time
##############################

# total_enroll_plot <- ggplot(total_males, aes(x = year, y = total)) +
#   geom_point(color = "black", alpha = 0.5, size = 1, shape = 15) +
#   geom_smooth(aes(year, total), method = "lm") +
#   #geom_line(color = "black") +
#   geom_point(data = total_females, color = "black", alpha = 0.5, size = 1, shape = 16) +
#   #geom_line(data = total_females, color = "black", linetype = "dotted") +
#   geom_smooth(aes(year, total), method = "lm") +
#   theme_classic() +
#   theme(panel.border = element_rect(colour = "black", fill = NA, size = 0.7)) +
#   labs(x = "Year", y = "Total Male & Female Graduate Student \nEnrollment in the U.S.") +
#   scale_y_continuous(labels = scales::scientific) 


total_enroll_plot <- ggplot(total, aes(x = year, y = total, shape = sex, color = sex)) +
  geom_point(size = 2, alpha = 0.5) +
  geom_smooth(method = "lm") +
  theme_classic() +
  theme(panel.border = element_rect(colour = "black", fill = NA, size = 0.7)) +
  labs(x = "Year", y = "Total Male & Female Graduate Student \nEnrollment in the U.S.") +
  scale_y_continuous(labels = scales::scientific) +
  scale_color_manual(values = c("grey11", "gray48")) +
  annotate("text", 
           x = 2010, 
           y = 940000, 
           label = "y = 9096x - 1.7e+07", 
           family = "Times New Roman", 
           size = 3) +
  annotate("text", 
           x = 2010, 
           y = 880000, 
           label = "italic(R)^2 == 0.85", 
           parse = TRUE, 
           family = "Times New Roman", 
           size = 3) +
  annotate("text", 
           x = 1999, 
           y = 1600000, 
           label = "y = 3.01e+04x - 5.9e+07", 
           family = "Times New Roman", 
           size = 3) +
  annotate("text", 
           x = 1999, 
           y = 1550000, 
           label = "italic(R)^2 == 0.98", 
           parse = TRUE, 
           family = "Times New Roman", 
           size = 3) 

total_enroll_plot
# Figure 1. Relationship between year and graduate student enrollment for both males and females (1967-2015). Year significantly predicts graduate student enrollment in the U.S. for both males (*b* = 9096, t(47) = 16.61, *p* <0.001) and females (*b* =  3.013e+04, t(47) = 51.66, *p* < 0.001) with a strong positive correlation between year and both sexes (Pearson's *r* (males) = 0.92, Pearson's *r* (females) =  0.99). The overall models (male enrollment = 9096(year) - 1.7e+07, female enrollment =  3.01e+04(year) - 5.9e+07) explain a significant amount of variance in salary for males (F(1,47) = 276, *p* < 0.001), R^2^ = 0.85 and females (F(1,47) = 2669, *p* < 0.001, R^2^ = 0.98). Gray region is the 95% confidence interval for the mean predicted values.

##############################
# fit a lm to male and female enrollment through time
##############################

# lm for males
enrollment_males_lm <- lm(total ~ year, data = total_males)
summary(enrollment_males_lm)
cor.test(total_males$total, total_males$year) # 0.92

# lm for females
enrollment_females_lm <- lm(total ~ year, data = total_females)
summary(enrollment_females_lm)
cor.test(total_females$total, total_females$year) # 0.99

##############################
# net & percent increase in total males
##############################

m_1967 <- total_males %>% # subset 1967 males
  filter(year == 1967)

m_2015 <- total_males %>% # subset 2015 males
  filter(year == 2015)

net_males <- m_2015$total - m_1967$total # net increase from 1967-2015
fold_males <- m_2015$total/m_1967$total # fold increase from 1967-2015
perc_males <- ((m_2015$total - m_1967$total)/m_1967$total) * 100 # % increase from 1967 - 2015

##############################
# net & percent increase in total females
##############################

f_1967 <- total_females %>% # subset 1967 females
  filter(year == 1967)

f_2015 <- total_females %>% # subset 2015 females 
  filter(year == 2015)

net_females <- f_2015$total - f_1967$total # net increase from 1967-2015
fold_females <- f_2015$total/f_1967$total # fold increase from 1967-2015
perc_females <- ((f_2015$total - f_1967$total)/f_1967$total) * 100 # % increase from 1967 - 2015

##############################
# findings
##############################

# CREATE TALE OF MALE & FEMALE RAW VALUES FOR 1967 & 2015 IN WORD DOC
```

Since 1967, there has been an overall increase in total (part-time, full-time, private, public) graduate student enrollment in the United States (Table 1, Fig. 1). Year is highly correlated with both male (Pearson's *r* = 0.92) and female (Pearson's *r* = 0.99) enrollemnt in graduate programs, though female enrollment has risen at a much faster rate (30,100 students/year) than males (9,096 students/year). As a result, females enrollment in 2015 was 6.5 times greater than in 1967, as compared to the 1.9 fold increase observed for males over the same period of time (Table 1). 

Things to research for discussion:
-social changes -- women away from being stay-at-home moms/housekeepers and towards being higher degree earners in the workforce
-why have then increased faster than males??

```{r shifts in female PhD recipients by field (1985, 2000, 2015)--Sam's plot and chi square}

#SOURCE: National Science Foundation, National Center for Science and Engineering Statistics, Survey of Earned Doctorates. You can find more information here: https:// www.nsf.gov/ statistics/ srvydoctorates/ .

##############################
# subset out females by field from years 1985, 2000, 2015
##############################

female_phd <- fields %>% 
  filter(field == "phys_earth_female" | 
           field == "engineering_female" | 
           field == "edu_female" | 
           field == "humanities_arts_female") %>% 
  select(field, `1985_raw`, `2000_raw`, `2015_raw`) %>% 
  rename(`1985` = `1985_raw`, `2000` = `2000_raw`, `2015` = `2015_raw`) 

##############################
# find total females across all four fields for each year and subset out each year's total
##############################

female_phd_total <- female_phd %>% 
  summarise(
    total_1985 = sum(`1985`),
    total_2000 = sum(`2000`),
    total_2015 = sum(`2015`)
  )

tot_1985 <- female_phd_total$total_1985
tot_2000 <- female_phd_total$total_2000
tot_2015 <- female_phd_total$total_2015

##############################
# add column to female_phd df with calculated percentages
##############################

female_phd_perc <- female_phd %>% 
  mutate(perc_1985  = round((`1985`/tot_1985) * 100, 1), perc_2000 = round((`2000`/tot_2000) * 100, 1), perc_2015 = round((`2015`/tot_2015) * 100, 1)) %>% # calculate percentage of females within each of the four fields
  select(field, perc_1985, perc_2000, perc_2015) %>% 
  mutate(
    field_long = case_when( # do this so legend labels are correct
      field == "edu_female" ~ "Education",
      field == "humanities_arts_female" ~ "Humanities & Arts",
      field == "engineering_female" ~ "Engineering",
      field == "phys_earth_female" ~ "Physical & Earth Sciences"
    )
  )

##############################
# convert to long format
##############################

female_phd_long <- melt(female_phd_perc, variable.name = "year", value.name = "percent") %>% 
 mutate(
   year_long = case_when(
     year == "perc_1985" ~ "1985",
     year == "perc_2000" ~ "2000",
     year == "perc_2015" ~ "2015"
   )
 ) %>% 
  select(field_long, year_long, percent)

##############################
# create stacked barplot
##############################

female_phd_plot <- ggplot(female_phd_long, aes(x = year_long, y = percent, fill = field_long)) + 
  geom_bar(stat = "identity", colour = "black") +
  geom_text(aes(x = year_long, y = percent, label = paste0(percent,"%")), size = 4,
            position = position_stack(vjust = 0.5)) +
  theme_classic() +
  theme(panel.border = element_rect(colour = "black", fill = NA, size = 0.7)) +
  scale_y_continuous(expand = c(0,0), limits = c(0,110)) +
  labs(x = "Year", y = "Female PhD Recipients by Field (%)") +
  scale_fill_manual(name = "Field of Study", values = c("snow4", "mediumaquamarine", "thistle4", "slategray2"))

female_phd_plot

# Figure 2: Percentage of female PhD recipients, by field, from 1985, 2000, & 2015. Percentages of female PhD recipients from 1985 (n = 5,650), 2000 (n = 8,711), and 2015 (n = 10,623) were assessed across 4 focal fields of study: Education, Engineering, Humanities & Arts, Physical & Earth Sciences. Fields assessed include education, Other academic fields have not been considered for this plot.... THIS STILL ISN'T RIGHT...

##############################
# wrangle data into format for eventual presentation
##############################

female_phd_chi2_present <- fields %>% 
  filter(field == "phys_earth_female" | # filter for desired fields of study
           field == "engineering_female" | 
           field == "edu_female" | 
           field == "humanities_arts_female") %>% 
  select(field, `1985_raw`, `2000_raw`, `2015_raw`) %>%  # select raw data from years of interest
  rename(`1985` = `1985_raw`, `2000` = `2000_raw`, `2015` = `2015_raw`) %>% # rename year variables 
  mutate(perc_1985  = round((`1985`/tot_1985) * 100, 1), perc_2000 = round((`2000`/tot_2000) * 100, 1), perc_2015 = round((`2015`/tot_2015) * 100, 1)) %>% # calculate percentage of females within each of the four fields
  select(field, `1985`, perc_1985, `2000`, perc_2000, `2015`, perc_2015)

# Table 2: PhDs awarded to female graduate students, by field and year. Values indicated are counts, with proportions (by field of study) in italics. Data source: ___.

##############################
# wrangle data into format for chi2 test
##############################

female_phd2 <- fields %>% 
  filter(field == "edu_female" | 
        field == "engineering_female" |
        field == "humanities_arts_female" |
        field == "phys_earth_female") %>% 
  select(`1985_raw`, `2000_raw`, `2015_raw`) %>% 
  rename(`1985` = `1985_raw`, `2000` = `2000_raw`, `2015` = `2015_raw`) 

rownames(female_phd2) <- c("Physical & Earth Sciences", "Engineering", "Education", "Humanities & Arts")

##############################
# chi2 test (use female_phd df from earlier section to run test)
##############################

# QUESTION: Is there a significant effect of year on the proportion of PhD type awarded to female graduate students?


x2_test <- chisq.test(female_phd2)

# FINDING: Field of study differs significantly for female PhD recipients in the years 1985, 2000, and 2015.
```

```{r An's chi square}
#### chi square for 1985 ####
fields_85 <- phd_fields_field %>%
  filter(year == "1985",
         field == "lifesci" |
           field == "physsci" |
           field == "math" |
           field == "psych" |
           field == "eng" |
           field == "edu" |
           field == "arts" |
           field == "other") %>%
  select("male", "female")

rownames(fields_85) <- c("lifesci", "physsci", "math", "psych", "eng", "edu", "arts", "other")

fields_85_prop <- phd_fields_field %>%
  filter(year == "1985",
         field == "lifesci" |
           field == "physsci" |
           field == "math" |
           field == "psych" |
           field == "eng" |
           field == "edu" |
           field == "arts" |
           field == "other") %>%
  select("male_prop", "female_prop")

rownames(fields_85_prop) <- c("lifesci", "physsci", "math", "psych", "eng", "edu", "arts", "other")

fields85_chi <- chisq.test(fields_85)
fields85_chi
# proportions are different, no shit

#### chi square for 2000 ####

fields_00 <- phd_fields_field %>%
  filter(year == "2000",
         field == "lifesci" |
           field == "physsci" |
           field == "math" |
           field == "psych" |
           field == "eng" |
           field == "edu" |
           field == "arts" |
           field == "other") %>%
  select("male", "female")

rownames(fields_00) <- c("lifesci", "physsci", "math", "psych", "eng", "edu", "arts", "other")

fields_00_prop <- phd_fields_field %>%
  filter(year == "2000",
         field == "lifesci" |
           field == "physsci" |
           field == "math" |
           field == "psych" |
           field == "eng" |
           field == "edu" |
           field == "arts" |
           field == "other") %>%
  select("male_prop", "female_prop")

rownames(fields_00_prop) <- c("lifesci", "physsci", "math", "psych", "eng", "edu", "arts", "other")

fields00_chi <- chisq.test(fields_00)
fields00_chi
# proportions are different

#### chi square for 2015 ####

fields_15 <- phd_fields_field %>%
  filter(year == "2015",
         field == "lifesci" |
           field == "physsci" |
           field == "math" |
           field == "psych" |
           field == "eng" |
           field == "edu" |
           field == "arts" |
           field == "other") %>%
  select("male", "female")

rownames(fields_15) <- c("lifesci", "physsci", "math", "psych", "eng", "edu", "arts", "other")

fields_15_prop <- phd_fields_field %>%
  filter(year == "2015",
         field == "lifesci" |
           field == "physsci" |
           field == "math" |
           field == "psych" |
           field == "eng" |
           field == "edu" |
           field == "arts" |
           field == "other") %>%
  select("male_prop", "female_prop")

rownames(fields_15_prop) <- c("lifesci", "physsci", "math", "psych", "eng", "edu", "arts", "other")

fields15_chi <- chisq.test(fields_15)
fields15_chi
#proportions are different
```

```{r specialty line graph} 
line_df <- phd_fields_field %>% 
  filter(year == "1985" |
           year == "2000" |
           year == "2015") %>%
  filter(field == "physsci" |
           field == "eng" |
           field == "edu" |
           field == "arts") %>% 
  select("field", "year", "female_prop")

physsci <- phd_fields_field %>% 
  filter(field == "physsci") %>% 
  select("year", "female_prop")

eng <- phd_fields_field %>% 
  filter(field == "eng") %>% 
  select("year", "female_prop")

edu <- phd_fields_field %>% 
  filter(field == "edu") %>% 
  select("year", "female_prop")

arts <- phd_fields_field %>% 
  filter(field == "arts") %>% 
  select("year", "female_prop")

line_x = "Year"
line_y = "Proportion of female students in total population"

fields_line <- ggplot(line_df, aes(x = year, y = female_prop, group = field, color = field)) +
  geom_line() +
  geom_point(size = 3) +
  scale_x_continuous(breaks = c(1985, 2000, 2015), labels = c("1985", "2000", "2015")) +
  scale_color_manual(values = beyonce_palette(18)) +
  expand_limits(y = 1) +
  theme_bw() +
  theme(panel.grid.minor = element_line(linetype = "blank")) +
  labs(x = line_x, y = line_y)
  
fields_line
```

```{r chi square again}
#### physsci ####
physsci_x2df <- phd_fields_field %>% 
  filter(field == "physsci") %>% 
  select("male", "female")

rownames(physsci_x2df) <- c("1985", "1990", "1995", "2000", "2005", "2010", "2015")

physsci_x2 <- chisq.test(physsci_x2df)
physsci_x2

#### eng ####
eng_x2df <- phd_fields_field %>% 
  filter(field == "eng") %>% 
  select("male", "female")

rownames(eng_x2df) <- c("1985", "1990", "1995", "2000", "2005", "2010", "2015")

eng_x2 <- chisq.test(eng_x2df)
eng_x2

#### edu ####
edu_x2df <- phd_fields_field %>% 
  filter(field == "edu") %>% 
  select("male", "female")

rownames(edu_x2df) <- c("1985", "1990", "1995", "2000", "2005", "2010", "2015")

edu_x2 <- chisq.test(edu_x2df)
edu_x2

#### arts ####
arts_x2df <- phd_fields_field %>% 
  filter(field == "arts") %>% 
  select("male", "female")

rownames(arts_x2df) <- c("1985", "1990", "1995", "2000", "2005", "2010", "2015")

arts_x2 <- chisq.test(arts_x2df)
arts_x2

```
Comparing total PhD students, there is a significant effect of year on proportion of female students in the physical sciences, (_X_^2^(`r physsci_x2$parameter`) = `r physsci_x2$statistic`, _p_ < 0.001), engineering (_X_^2^(`r eng_x2$parameter`) = `r eng_x2$statistic`, _p_ < 0.001), education (_X_^2^(`r edu_x2$parameter`) = `r edu_x2$statistic`, _p_ < 0.001), and the humanities and arts (_X_^2^(`r arts_x2$parameter`) = `r arts_x2$statistic`, _p_ < 0.001). 
Between 1985 and 2015, female PhD students have shown a growing interest in the physical sciences and engineering, as the percent of female students specializing in the physical sciences has increased 8.6% from 10.1% to 18.7% (cite stacked bar graph) and the percent of female students specializing in engineering has increased 18.2% since 1985. This is reflected in the growing proportion of female students in the total number of students in the physical sciences has doubled between 1985 (0.16) and 2015 (0.34) and quadrupled between 1985 (0.06) and 2015 (0.23) in engineering.

```{r male & female salaries for starting postdoc and other employment positions (2015)}

# SOURCE: National Science Foundation, National Center for Science and Engineering Statistics, Survey of Earned Doctorates, 2015.

##############################
# make data into long format
##############################

postdoc_salary_long <- melt(postdoc_salary, variable.name = "employment_status", value.name = "salary") 

##############################
# add column with "sex"
##############################

postdoc_salary_new <- postdoc_salary_long %>%
  mutate(
    sex = case_when(
      employment_status == "employment_males" ~ "Male",
      employment_status == "postdoc_study_males" ~ "Male",
      employment_status == "employment_females" ~ "Female",
      employment_status == "postdoc_study_females" ~ "Female"
    )
  ) %>% 
  mutate(
    status = case_when(
      employment_status == "employment_males" ~ "Employed",
      employment_status == "postdoc_study_males" ~ "Postdoc-study",
      employment_status == "employment_females" ~ "Employed",
      employment_status == "postdoc_study_females" ~ "Postdoc-study"
    )
  )

##############################
# make status a factor, and list levels in the desired facet order
##############################

postdoc_salary_new$status<- factor(postdoc_salary_new$status, levels=c("Postdoc-study", "Employed"))

##############################
# boxplot of employment males vs. females & postdoc males vs. females
##############################

postdoc_salary_plot <- ggplot(postdoc_salary_new, aes(x = sex, y = salary)) +
  geom_boxplot(aes(group = sex, fill = factor(sex))) +
  scale_fill_manual(values = c("white", "darkgrey"), name = "Sex") +
  facet_grid(~ status) +
  labs(x = "", y = "Salary (USD)") +
  scale_y_continuous(label = dollar_format()) +
  theme_classic() +
  theme(
    axis.text.x = element_blank(),
    axis.ticks.x=element_blank(),
    panel.border = element_rect(colour = "black", fill=NA, size=0.7))

postdoc_salary_plot

# Figure 3: Median annual salary (USD) for doctorate recipients with definite postgraduations plans in the U.S., by sex (2015). Median salary does not significantly differ between males and females, regardless of postgraduation plans. Data source: ____.

#mann-whitney u?

##############################
# separately subset males & females by employed & postdoc status
##############################

male_postdoc <- postdoc_salary_new %>% 
  filter(sex == "Male", status == "Postdoc-study")

female_postdoc <- postdoc_salary_new %>% 
  filter(sex == "Female", status == "Postdoc-study")

male_employed <- postdoc_salary_new %>% 
  filter(sex == "Male", status == "Employed")

female_employed <- postdoc_salary_new %>% 
  filter(sex == "Female", status == "Employed")

##############################
# data is not continuous (pre-calculated median salaries by sex and field), so instead run a nonparametric test comparing salaries between males and females for postdoc and employment positions
  # Advantages of non-parametric tests:
    # no assumption of normality or equal variance
    # compares medians (which is actually a better parameter for income comparisons)
# Use a Mann-Whitney U test (non-parametric alternative for unparied t-test)
  # H0: ranks are equal
  # Ha: ranks are unequal
##############################

postdoc_u <- wilcox.test(male_postdoc$salary, female_postdoc$salary) # p = 0.87
# Median salary does not significantly differ between males and females starting postdoc positions in 2015.

employed_u <- wilcox.test(male_employed$salary, female_employed$salary) # p = 0.33
# Median salary does not significantly differ between males and females PhD recipients in non-postdoc employment positions.

##############################
# effect size
##############################

postdoc_d <- cohen.d(male_postdoc$salary, female_postdoc$salary) # negligible
abs_postdoc <- median(male_postdoc$salary) - median(female_postdoc$salary) # males > females by $3000
perc_postdoc <- (median(male_postdoc$salary) - median(female_postdoc$salary))/(median(female_postdoc$salary)) * 100 # male postdoc salary 6.7% higher than females

employed_d <- cohen.d(male_employed$salary, female_employed$salary) # small
abs_employed <- median(male_employed$salary) - median(female_employed$salary) # males > females by $3417
perc_employed <- (median(male_employed$salary) - median(female_employed$salary))/(median(female_employed$salary)) * 100  # male employed salary 4.8% higher than females
```

Discussion Points:
- despite non-significant differences, males still make more money than females in both groups
- discuss wage gap

```{r exploring academic salaries for profs in US colleges}


##### Shifts in female PhD recipients by field (1985, 2000, and 2015). 
#Describe if and how there was a shift in PhDs awarded to females in four fields (Physical and Earth Sciences, Engineering, Education, and Humanities & Arts) in 1985, 2000, and 2015. Describe your results statistically, in a graph or table, and in text. Note: There are several ways that you can interpret this question. You are invited to decide which you think is/are most interesting. Just be really clear about what you are asking/answering in your report.

```

##### Model describing faculty salary

Explore relationships between variables in the ‘Faculty salary data (2008 - 2009 survey)’ dataset. Develop a model describing faculty salary based on data for faculty sex, rank, years in current position, field, and number of years since doctoral degree was earned. You should make decisions regarding which variables should remain in your final model. Describe the results qualitatively and quantitatively (i.e., don’t just report the statistical results of the model – make sure you describe interesting findings in text). You can also discuss any concerns that you have with the model(s) you present, if any.

```{r visualize faculty_salary data}
##############################
# Look at data
##############################

fac_by_dept_plot <- ggplot(data = faculty_salary, aes(x = years_service, y = salary)) +
  geom_point(aes(color = rank, pch = sex)) +
  facet_wrap(~ discipline)
# too much variability to determine a stat. sig effect between sexes? see models below

##############################
# Explore relationships between variables that could be correlated
##############################

# looking for colinearity between years_since and years_service
years_plot <- ggplot(faculty_salary, aes(x = years_since, y = years_service)) +
  geom_point()
# FINDINGS: appear to be correlated (i.e. increase in years_since = ~ linear increase in years_service)
```

```{r multivariate linear regression for faculty salary}

# run full model then find VIF to look for colinearity
full_lm1 <- lm(salary ~ rank + discipline + years_since + years_service + sex, data = faculty_salary)
summary(full_lm1) # R2 = 0.44, p < 0.002
plot(full_lm1) # does NOT look good. Heteroscedastic??
vif(full_lm1)
# FINDINGS: years_since & years_service both have VIF > 4 (some correlation, think critically about possible overlaps); this in combination with "years_plot" make me think that we should drop one of these predictor variables--I'm inclined to keep years_service because I think it makes more sense that the longer that you've been faculty, the higher your rank (see fac_by_dept_plot)--which typically means the higher your salary

# lm 2 includes rank, discipline, sex, years_serive (eliminated years_since becuase it appeared correlated with years_service)
facsal_lm2 <- lm(salary ~ rank + discipline + years_service + sex, data = faculty_salary)
summary(facsal_lm2) # R2 = 0.44, p < 0.001 # reference = female, A, Associate Prof
plot(facsal_lm2) # don't look any better than lm1
vif(facsal_lm2) # no correlataion between variables GOOD!
# FINDINGS: not necessarily non-linear (residuals vs fitted); appears to be heteroscedastic (Scale-Location plot)

# lm3 includes discipline, sex, years_service (eliminated rank because rank and years_service should be correlated and may be competing to explain variation...hence why years_service is not sig in lm2)
facsal_lm3 <- lm(salary ~ discipline + years_service + sex, data = faculty_salary)
summary(facsal_lm3) # R2 = 0.16...NOT GOOD, p < 0.001 # reference = female, A
plot(facsal_lm3) # these look A LOT BETTER than lm1, lm2!!! no heteroscedasticity
vif(facsal_lm3) # no correlation
```

```{r AIC values}
#### all disciplines ####
AIC_full_lm1 <- AIC(full_lm1)
AIC_full_lm1
# 9093.826

AIC_lm2 <- AIC(facsal_lm2) # difference of 2 in AIC may be considered not different...this is pretty close to lm1 so I'd be inclined to keep this over the full model
AIC_lm2
# 9096.813

AIC_lm3 <- AIC(facsal_lm3) # WELL FUCK. definitely not good, thought i may have been on to something...
AIC_lm3
# 9257.162
```

###References:

(Figure 1) U.S. Department of Education, National Center for Education Statistics, Higher Education General Information Survey (HEGIS), "Fall Enrollment in Colleges and Universities" surveys, 1967 through 1985; Integrated Postsecondary Education Data System (IPEDS), "Fall Enrollment Survey" (IPEDS-EF:86-99); IPEDS Spring 2001 through Spring 2016, Fall Enrollment component; and Enrollment in Degree-Granting Institutions Projection Model, 2000 through 2026. Accessed at: https:// nces.ed.gov/ programs/digest/ d16/ tables/ dt16_303.80.asp.

(Figure 2) National Science Foundation, National Center for Science and Engineering Statistics, Survey of Earned Doctorates. https:// www.nsf.gov/ statistics/ srvydoctorates/ .

(Figure 3) National Science Foundation, National Center for Science and Engineering Statistics, Survey of Earned Doctorates, 2015.

* differences in faculty salaries based on discipline: https://www.insidehighered.com/news/2016/03/28/study-finds-continued-large-gaps-faculty-salaries-based-discipline
* maybe a good discussion point, even though public vs. private not a variable we consider in our data set: http://www.nea.org/home/34420.htm