---
title: "Trends in Graduate Enrollment, Shifting Fields of Study, and Faculty Salary"
author: "Sam Csik, An Bui"
date: "5 December 2018"
output: word_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

**Introduction: ** 
-	Women have improved access to post-baccalaureate education  
-	However, there is much gender disparity in discipline-specific involvement and salaries  
-	In the US specifically, faculty salary has increased but female faculty consistently earn less than male faculty (cite table, https://nces.ed.gov/programs/digest/d17/tables/dt17_316.20.asp) 

*Objectives* (we should put this in intro this time):  
  The following report uses publicly available demographic data to evaluate changes in the types of advanced academic degrees sought and by whom, as well as salaries of doctorate degree earners both inside and outside academia and at different stages of their careers. We use these data to assess: 
-	trends in the gender distribution of doctoral students in the US from 1985-2015,   
-	shifts in female enrollment in doctoral programs in the Physical Sciences, Engineering, Education, and the Humanities and Arts in two ways:  
    - how female graduate student research interests have shifted between these four fields, and  
    - how the proportion of female graduate students in these fields have changed  
-	differences in salaries, by sex, after receiving a PhD 
-	variables that may predict faculty salaries.

**Data & Methods:**
  All data were analyzed using R version 3.5.1. We first investigated changes in graduate student enrollment in the United States using data provided by the National Center for Education Statistics (NCES) (cite). We modeled the relationship between year and male and female graduate student enrollment between 1967 and 2015 by fitting linear equations to data on each sex and comparing the rate of change between the two groups.
  Using data provided by the National Science Foundation's (NSF) Survey of Earned Doctorates (SED) (cite), we assessed changes in female graduate student research interests across four fields of study--education, engineering, humanities & arts, and the physical & earth sciences--between 1985, 2000, and 2015. A chi-squared test was used to determine whether there was a significant association between year and the type of PhDs awarded to women to evalue a shift in research interests through time. ADD AN'S ANALYSIS HERE. 
  To determine if there are presently (2015) pay disparities between men and women with graduate degrees, we then compared salaries between the sexes in both postdoc and non-postdoc employment positions using a Mann-Whitney U test ($alpha$ = 0.05). Data was again provided by NSF's SED.
  We lastly constructed a multivariate linear model to describe faculty salary based on X, Y, Z. We later suggest additional parameters that we believe may help to better inform our model.
 
**Analysis & Discussion**
```{r load packages and data, include = FALSE}

library(tidyverse)
library(reshape2) 
library(effsize)
library(scales)
library(extrafont)
library(car)
library(stargazer)
library(beyonce)

enrollment <- read_csv("grad_enrollment_copy.csv")
postdoc_salary <- read_csv("median_salary_copy.csv")
fields <- read_csv("phds_by_field_copy.csv")
faculty_salary <- read_csv("faculty_salary.csv")
phd_fields_field <- read_csv("phds_field_field.csv")
```

##### Trends in gender distribution of doctoral students
```{r total enrollment plot}
#### subset total males and females ####
total_males <- enrollment %>% 
  select(year, total_males) %>% 
  mutate(total = total_males) %>% # add column called "total" that is equal to the "total_males" column so that we can plot males and females on same plot with same y-axis 
  mutate(sex = rep("male", length(total_males))) %>% 
  select(year, total, sex)

total_females <- enrollment %>% 
  select(year, total_females) %>% 
  mutate(total = total_females) %>%  # add column called "total" that is equal to the "total_females" column so that we can plot males and females on same plot with same y-axis
  mutate(sex = rep("female", length(total_females))) %>% 
  select(year, total, sex)

total <- rbind(total_males, total_females)

#### plot trends through time ####
total_enroll_plot <- ggplot(total, aes(x = year, y = total, shape = sex, color = sex)) +
  geom_point(size = 2, alpha = 0.5) +
  geom_smooth(method = "lm") +
  theme_classic() +
  theme(panel.border = element_rect(colour = "black", fill = NA, size = 0.7)) +
  labs(x = "Year", y = "Total Male & Female Graduate Student \nEnrollment in the U.S.") +
  scale_y_continuous(labels = scales::scientific) +
  scale_color_manual(values = c("grey11", "gray48")) +
  annotate("text", 
           x = 2010, 
           y = 940000, 
           label = "y = 9096x - 1.7e+07", 
           family = "Times New Roman", 
           size = 3) +
  annotate("text", 
           x = 2010, 
           y = 880000, 
           label = "italic(R)^2 == 0.85", 
           parse = TRUE, 
           family = "Times New Roman", 
           size = 3) +
  annotate("text", 
           x = 1999, 
           y = 1600000, 
           label = "y = 3.01e+04x - 5.9e+07", 
           family = "Times New Roman", 
           size = 3) +
  annotate("text", 
           x = 1999, 
           y = 1550000, 
           label = "italic(R)^2 == 0.98", 
           parse = TRUE, 
           family = "Times New Roman", 
           size = 3) 

total_enroll_plot
# Figure 1. Relationship between year and graduate student enrollment for both males and females (1967-2015). Year significantly predicts graduate student enrollment in the U.S. for both males (*b* = 9096, t(47) = 16.61, *p* <0.001) and females (*b* =  3.013e+04, t(47) = 51.66, *p* < 0.001) with a strong positive correlation between year and both sexes (Pearson's *r* (males) = 0.92, Pearson's *r* (females) =  0.99). The overall models (male enrollment = 9096(year) - 1.7e+07, female enrollment =  3.01e+04(year) - 5.9e+07) explain a significant amount of variance in salary for males (F(1,47) = 276, *p* < 0.001), R^2^ = 0.85 and females (F(1,47) = 2669, *p* < 0.001, R^2^ = 0.98). Gray region is the 95% confidence interval for the mean predicted values.
```
 *Male and female graduate student enrollment in the U.S. (1967-2015)*
  Since 1967, there has been an overall increase in total (part-time, full-time, private, public) graduate student enrollment in the United States (Table 1, Fig. 1). Year is highly correlated with both male (Pearson's *r* = 0.92) and female (Pearson's *r* = 0.99) enrollemnt in graduate programs, though female enrollment has risen at a much faster rate (30,100 students/year) than males (9,096 students/year). As a result, females enrollment in 2015 was 6.5 times greater than in 1967, as compared to the 1.9 fold increase observed for males over the same period of time (Table 1). 

Things to research for discussion:
-social changes -- women away from being stay-at-home moms/housekeepers and towards being higher degree earners in the workforce
-why have then increased faster than males??

##### Female graduate enrollment in Physical Sciences, Engineering, Education, and the Humanities and Arts
```{r stacked bar chart}
#### subset out females by field from years 1985, 2000, 2015 ####
female_phd <- fields %>% 
  filter(field == "phys_earth_female" | 
           field == "engineering_female" | 
           field == "edu_female" | 
           field == "humanities_arts_female") %>% 
  select(field, `1985_raw`, `2000_raw`, `2015_raw`) %>% 
  rename(`1985` = `1985_raw`, `2000` = `2000_raw`, `2015` = `2015_raw`) 


#### find total females across all four fields for each year and subset out each year's total ####
female_phd_total <- female_phd %>% 
  summarise(
    total_1985 = sum(`1985`),
    total_2000 = sum(`2000`),
    total_2015 = sum(`2015`)
  )

tot_1985 <- female_phd_total$total_1985
tot_2000 <- female_phd_total$total_2000
tot_2015 <- female_phd_total$total_2015

#### add column to female_phd df with calculated percentages ####
female_phd_perc <- female_phd %>% 
  mutate(perc_1985  = round((`1985`/tot_1985) * 100, 1), perc_2000 = round((`2000`/tot_2000) * 100, 1), perc_2015 = round((`2015`/tot_2015) * 100, 1)) %>% # calculate percentage of females within each of the four fields
  select(field, perc_1985, perc_2000, perc_2015) %>% 
  mutate(
    field_long = case_when( # do this so legend labels are correct
      field == "edu_female" ~ "Education",
      field == "humanities_arts_female" ~ "Humanities & Arts",
      field == "engineering_female" ~ "Engineering",
      field == "phys_earth_female" ~ "Physical & Earth Sciences"
    )
  )

#### convert to long format ####
female_phd_long <- melt(female_phd_perc, variable.name = "year", value.name = "percent") %>% 
 mutate(
   year_long = case_when(
     year == "perc_1985" ~ "1985",
     year == "perc_2000" ~ "2000",
     year == "perc_2015" ~ "2015"
   )
 ) %>% 
  select(field_long, year_long, percent)

#### create stacked barplot ####
female_phd_plot <- ggplot(female_phd_long, aes(x = year_long, y = percent, fill = field_long)) + 
  geom_bar(stat = "identity", colour = "black") +
  geom_text(aes(x = year_long, y = percent, label = paste0(percent,"%")), size = 4,
            position = position_stack(vjust = 0.5)) +
  theme_classic() +
  theme(panel.border = element_rect(colour = "black", fill = NA, size = 0.7)) +
  scale_y_continuous(expand = c(0,0), limits = c(0,110)) +
  labs(x = "Year", y = "Female PhD Recipients by Field (%)") +
  scale_fill_manual(name = "Field of Study", values = c("snow4", "mediumaquamarine", "thistle4", "slategray2"))

female_phd_plot
```
  Between 1985 and 2015, female interests in the four focal fields (Physical Sciences, Engineering, Education, and the Humanities and Arts) have shifted. Between 1985 and 2015, the proportion of female students specializing in education decreased by nearly half but the proportion of female students in engineering has increased seven-fold (Fig. X).  
- says something about increased access of women to engineering.
  
*Shifts in female PhD recipients by field (1985, 2000, 2015)*
  Comparing total PhD students, there is a significant effect of year on proportion of female students in the physical sciences, (_X_^2^(`r physsci_x2$parameter`) = `r physsci_x2$statistic`, _p_ < 0.001), engineering (_X_^2^(`r eng_x2$parameter`) = `r eng_x2$statistic`, _p_ < 0.001), education (_X_^2^(`r edu_x2$parameter`) = `r edu_x2$statistic`, _p_ < 0.001), and the humanities and arts (_X_^2^(`r arts_x2$parameter`) = `r arts_x2$statistic`, _p_ < 0.001). 
  To determine if the composition of female PhD recipients has shifted thorugh time, we assessed the proportions of women awarded PhDs across these same four focal fields of study. We found that the types of PhDs awarded differ significantly for female recipients in the years 1985, 2000, and 2015 ($\chi^2$(`r female_phd2$parameter`) = `r female_phd2$statistic`, *p* < 0.001). There has been a growing shift towards STEM fields (engineering and the physical & earth sciences) through time (Fig. stacked bar plot). While the majority of women received their advanced degrees in either education or the humanities & arts (combined total = 86.4%) in 1985, we see a steady diversification of interests across years with more equal representation across fields in the present day female graduate populace.
...(NOT SURE WHERE TO PUT THIS YET) This is reflected in the growing proportion of female students in the total number of students in the physical sciences has doubled between 1985 (0.16) and 2015 (0.34) and quadrupled between 1985 (0.06) and 2015 (0.23) in engineering.


*Salary differences between male and female PhD recipients*


*Modeling faculty salary*

**Conclusion:**
-	Female involvement in higher education has improved, but there are still steps to be made for progress  
    - While the proportion of female doctoral recipients has increased, only 29% of researchers worldwide are women (cite UNESCO, https://www.tellmaps.com/uis/gender/?lang=en#!/tellmap/79054752)
    - Female candidates for jobs are assessed based on whether or not they have children, regardless of the legality of this assessment  
-  While the data we have are as much as we can gather, there are missing components that would be useful for parameterizing a model describing faculty salaries
    - Institution type: research institutions vs liberal arts colleges  
    - Race: women make less than men, but Black women, Asian women, and Hispanic women make less than white women (there are numbers, will find later)  
-	Race is a particularly egregious variable to be left out, especially since it stratifies so much of higher education  
    - 64% of Bachelor’s degrees conferred are to white students (cite table, https://nces.ed.gov/programs/digest/d17/tables/dt17_322.20.asp)  
    -	66% of PhD recipients are white (cite table, https://nces.ed.gov/programs/digest/d17/tables/dt17_324.20.asp)  
    -	82% of professors are white (55% men, 27% women, cite figure, https://nces.ed.gov/fastfacts/display.asp?id=61)  
